---
title: "Shiny: NMES1988"
runtime: shiny
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
library(AER)
library(tidyverse)
data("NMES1988")

NMES1988 <- NMES1988 |>
    sample_n(500)    # use smaller data so shiny app runs faster
```


First create shiny app to color points by Private Insurance or not.  Note the use of `if` and `else if` to make use of the `checkboxInput`.

```{r, echo=FALSE, warning=FALSE}
inputPanel(
  checkboxInput("type_check", "Include private insurance?", value = FALSE)
)

renderPlot({
     if (input$type_check) 
        ggplot(NMES1988, aes(x = age, y = chronic))  + 
          geom_jitter(aes(color = insurance), size = 0.9, alpha = 0.4) +
          scale_color_manual(values=c("orange","blue"))
     else if (!input$type_check)      
        ggplot(NMES1988, aes(x = age, y = chronic))  + 
          geom_jitter(size = 0.9, alpha = 0.4) +
          scale_color_manual(values=c("orange","blue"))
})
```

Next create a shiny app that allows you to choose explanatory variables.  Note that we have to use enclose input variable names inside .data[[.]]; inputs are saved as strings (with quotes) but the tidyverse uses tidy evaluation (no quotes), so essentially the .data[[.]] strips off the quotes so that the variable can be recognized in the tidyverse.

```{r, echo=FALSE}
titlePanel("Predictors of physician visits in NMES1988 data")

inputPanel(
  selectInput("x", label = "x-axis variable:",
    choices = c("Family income" = "income",
                "Age (years/10)" = "age",
                "Chronic conditions" = "chronic")),
  checkboxInput("type_check2","Include private insurance?",value=FALSE)
)

renderPlot({
     if (input$type_check2) 
        ggplot(NMES1988, aes(x = .data[[input$x]], y = visits, 
                                    color = insurance)) +   
          scale_color_manual(values=c("orange","blue")) + 
          geom_jitter(size = 0.9, alpha = 0.4) 
     else if (!input$type_check2)      
        ggplot(NMES1988, aes(x = .data[[input$x]], y = visits)) +   
          scale_color_manual(values=c("orange","blue")) + 
          geom_jitter(size = 0.9, alpha = 0.4)
})
```

Try shiny app with a slider.  Slider inputs are treated as numeric.

```{r eruptions, echo=FALSE}
inputPanel(
  sliderInput("size", label = "Point size",
              min = 0.2, max = 2, value = 0.9, step = 0.1),
  sliderInput("alpha", label = "Point transparency",
              min = 0.01, max = 1, value = 0.4, step = 0.1),
  checkboxInput("type_check4","Include private insurance?",value=FALSE)
)

renderPlot({
     if (input$type_check4) 
        ggplot(NMES1988, aes(x = age, y = chronic))  + 
          geom_jitter(aes(color = insurance), 
                          size = input$size, alpha = input$alpha) +
          scale_color_manual(values=c("orange","blue"))
     else if (!input$type_check4)      
        ggplot(NMES1988, aes(x = age, y = chronic))  + 
          geom_jitter(size = input$size, alpha = input$alpha) +
          scale_color_manual(values=c("orange","blue"))
})

```


One more example featuring a boxplot:

```{r, echo = FALSE}
inputPanel(
  selectInput("boxplotx", label = "x-axis variable:",
    choices = c("Marital Status" = "married",
                "Health Status" = "health",
                "Private Insurance" = "insurance"))
)



renderPlot({
  ggplot(NMES1988, aes(y = visits, x = .data[[input$boxplotx]] )) + 
    geom_boxplot()
})
```


```{r, echo = FALSE, include = FALSE}
#Data Cleaning!
ic <- read_csv("data/ic2023.csv") |>
  select(college_id = UNITID, 
         study_abroad = SLO6, 
         room_capacity = ROOMCAP, 
         ncaa = ASSOC1, 
         college_ran_housing = ROOM, 
         meals_per_week = MEALSWK) 
  
ic_ay <- read_csv("data/ic2023_ay.csv") |>
  select(college_id = UNITID,
         in_state_tuition = TUITION2,
         food_housing_fee = CHG5AY3) 

hd <- read_csv("data/hd2023 2.csv") |>
  select(college_id = UNITID,
         college_name = INSTNM,
         city = CITY,
         state = STABBR,
         open_or_closed = CYACTIVE,
         size_type = INSTSIZE,
         longitude = LONGITUD,
         latitude = LATITUDE,
         public_private = CONTROL)

effy <- read_csv("data/effy2023.csv") |> 
  select(college_id = UNITID,
         student_type = EFFYLEV,
         total_enrollment = EFYTOTLT,
         total_men = EFYTOTLM,
         total_women = EFYTOTLW) |>
  filter(student_type == "2") #filters for undergraduate enrollment numbers 

s <- read_csv("data/s2023_sis.csv") |>
  select(college_id = UNITID,
         employment_type = FACSTAT, 
         total_faculty = SISTOTL) |>
  filter(employment_type == "10") |> #staff with faculty status
  select(-employment_type)
```

```{r, echo = FALSE, include = FALSE}
#Joining clean data sets
ic_full <- ic |>
  left_join(ic_ay)  

ic_hd <- ic_full |>
  left_join(hd)

college_data <- ic_hd |>
  left_join(effy)

college_data2 <- college_data |>
  left_join(s)

mn_college_data <- college_data2 |>
  filter(state == "MN") |>
  arrange(college_name)
```

```{r}
inputPanel(
  selectInput("public_private", label = "Public/Private",
    choices = c("Public" = "Public",
                "Private ((not-for-profit)" = "Private (not-for-profit)", 
                "Private (for-profit)" = "Private (for-profit)"))
)

renderTable({
  mn_college_data |>
  select(college_name, public_private) |>
  filter(public_private != "NA") |>
  mutate(public_private = as.factor(public_private),
         public_private = fct_recode(public_private,
                                     "Public" = "1",
                                     "Private (not-for-profit)" = "2",
                                     "Private (for-profit)" = "3")) |>
    filter(college_type == input$public_private)
  rename(college_name, "College Name" = "college_name",
         public_private, "College Type" = "public_private") |>
  kable()
})

```

 
```