---
title: "01 Mini Project: Interactive Maps"
author: "Solveig Senf"
sidebar: false
format:
  html: default
editor_options: 
  chunk_output_type: console
---

This project features two maps (a static and interactive version) that display data from the Centers for Disease Control and Prevention (CDC). Data on Covid-19 levels was collected from February 2022 to May 2023 on a county level. This project displays data recorded on May 11, 2023. For the purposes of this state-level project, I will use the average of county-level data for each state to display the new Covid-19 cases per 100k within the previous 7 days. Data can be accessed [here](https://data.cdc.gov/Public-Health-Surveillance/United-States-COVID-19-Community-Levels-by-County/3nnm-4jni/about_data). 

Additionally, this project contains two maps (a static and interactive version) that display the 2024 Presidential Election results. Election data is from the Federal Election Commission (FEC) and can be found at [this link](https://www.fec.gov/introduction-campaign-finance/election-results-and-voting-information/). 

```{r}
library(tidyverse)
library(mdsr)
library(maps)
library(viridis)
library(lubridate)
library(leaflet)
library(sf)
library(RColorBrewer)
```

COVID-19 Data

Interactive Map #1

```{r}
#data set from the CDC
cdc_data <- read.csv("~/Downloads/01_United_States_COVID-19_Community_Levels_by_County_20250216.csv")
```

```{r}
covid_data <- cdc_data |>
  filter(date_updated == "2023-05-11", 
         !(state %in% c("Puerto Rico", "American Samoa", "Commonwealth of the Northern Mariana Islands", "United States Virgin Islands", "Guam")))
```

```{r}
states_polygon <- as_tibble(map_data("state")) |>
  select(region, group, order, lat, long)

states_sf <- read_sf("https://rstudio.github.io/leaflet/json/us-states.geojson") |>
  select(name, geometry)
```

```{r}
#convert county-level data to state-level data
covid_state_level_data <- covid_data |>
  filter(covid_cases_per_100k != "NA") |>
  group_by(state) |>
  summarize(covid_cases_per_100k = mean(covid_cases_per_100k))
```

```{r}
#format state names in all data sets so they match 
covid_state_level_data <- covid_state_level_data |>
  mutate(state = str_to_lower(state),
         state = str_replace_all(state, " ", ""),
         state = str_squish(state))

states_sf <- states_sf |>
  mutate(name = str_to_lower(name),
         name = str_replace_all(name, " ", ""))

states_polygon <- states_polygon|>
  mutate(region = str_replace_all(region, " ", ""))
```

```{r}
#join covid data with map data
covid_map <- covid_state_level_data |>
  right_join(states_polygon, by =c("state" = "region"))

covid_map <- covid_map |>
  right_join(states_sf, by =c("state" = "name"))
```

```{r}
states <- read_sf("https://rstudio.github.io/leaflet/json/us-states.geojson") 
class(states)

states <- states |>
  mutate(name = str_to_lower(name),
         name = str_replace_all(name, " ", ""))
```

```{r}
#join covid data with state plotting data
state_plotting_covid <- covid_state_level_data |>
  right_join(states, by = c("state" = "name"))

#bins for interactive map
state_plotting_covid_sf <- state_plotting_covid |>
  mutate(covid_intervals = cut(covid_cases_per_100k, n = 8,
          breaks = c(0, 10, 20, 30, 40, 50, 60, 70, Inf))) |>
  filter(!(state %in% c("alaska", "hawaii")))

#base map for interactive map
ggplot(data = state_plotting_covid_sf) + 
  geom_sf(aes(geometry = geometry, fill = covid_intervals), colour = "white", linetype = 2) + 
  theme_void() +  
  scale_fill_brewer(palette = "Blues") 
```

```{r}
bins <- c(0, 10, 20, 30, 40, 50, 60, 70, Inf)
pal <- colorBin("Blues", domain = state_plotting_covid_sf$covid_cases_per_100k, bins = bins)

library(htmltools)
library(glue)

str(state_plotting_covid_sf$covid_cases_per_100k)

#add labels to the state_plotting_covid_sf
state_plotting_sf_labels <- state_plotting_covid_sf |>
  mutate(labels = str_c(state, ": ", format(round(covid_cases_per_100k, 1)), " cases / 100k people"))

labels <- lapply(state_plotting_sf_labels$labels, HTML)

state_plotting_covid_sf <- st_as_sf(state_plotting_covid_sf)

leaflet(state_plotting_covid_sf) |>
  setView(-100, 42, 3.5) |> #to view the united states
  addTiles() |>
  addPolygons(
    fillColor = ~pal(covid_cases_per_100k),
    weight = 1,
    opacity = 2,
    color = "black",
    dashArray = "2",
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 2,
      color = "#FFFFFF",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) |>
  addLegend(pal = pal, values = ~covid_cases_per_100k, opacity = 0.7, title = "COVID-19 Cases per 100k People",
    position = "topright")
```

Election Data

Interactive Map #2

```{r}
#data from the federal election commission
election_data <- read.csv("~/Downloads/2024presgeresults.csv")

#tidy data so it can be used to plot
election_data <- election_data |>
  select(STATE, ELECTORAL.VOTE..TRUMP..R., ELECTORAL.VOTE..HARRIS..D., HARRIS, TRUMP) |>
  rename(state = STATE, 
         Trump = ELECTORAL.VOTE..TRUMP..R., 
         Harris = ELECTORAL.VOTE..HARRIS..D.,
         votes_harris = HARRIS,
         votes_trump = TRUMP) |>
  pivot_longer(cols = c(Trump, Harris),
               names_to ="candidate_won",
               values_to = "electoral_votes") |>
  filter(electoral_votes != "NA") |> #remove observations that show the loosing candidate in each state
  slice(-30, -20) #remove rows 30 and 20 which are second observations of maine and nebraska due to the fact that they split electoral votes. the observation recording the winning candidate is kept. 

election_data <- election_data |>
  mutate(state = str_to_lower(state),
         state = str_replace_all(state, " ", ""),
         state = str_squish(state)) |>
  mutate(votes_harris = str_squish(votes_harris),
         votes_trump = str_squish(votes_trump))
```

```{r}
electoral_map <- election_data |>
  right_join(states_polygon, by =c("state" = "region"))

electoral_map <- electoral_map |>
  right_join(states_sf, by =c("state" = "name"))
```

```{r}
election_map_sf <- election_data |>
  right_join(states, by = c("state" = "name"))

ggplot(data = election_map_sf) + 
  geom_sf(aes(geometry = geometry, fill = candidate_won), colour = "white", linetype = 2) + 
  scale_fill_manual(values = c("Harris" = "darkblue", "Trump" = "darkred")) + 
  labs(fill = "Winning Candidate") +
  theme_void() 
```

```{r}
pal <- colorFactor(c("darkblue", "darkred"), domain = election_map_sf$candidate_won) #colorFactor used for categorical variables

library(htmltools)
library(glue)

str(election_map_sf$votes_harris, votes_trump)

#add labels to the election_map_sf data set
election_map_sf <- election_map_sf |>
  mutate(labels = str_c(state, ": ", format(votes_harris), " votes Harris", ", ", format(votes_trump), " votes Trump"))

labels <- lapply(election_map_sf$labels, HTML)

election_map_sf <- st_as_sf(election_map_sf)

leaflet(election_map_sf) |>
  setView(-100, 42, 3.5) |> #to view the united states
  addTiles() |>
  addPolygons(
    fillColor = ~pal(candidate_won),
    weight = 1,
    opacity = 2,
    color = "black",
    dashArray = "2",
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 2,
      color = "#FFFFFF",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) |>
  addLegend(pal = pal, values = ~candidate_won, opacity = 0.7, title = "Winning Presidential Candidate",
    position = "topright")
```

